---
title: "Generliased Linear Model: Bootstrapping"
subtitle: "Week 17"
author: "Winnie Xia"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
class: inverse, center, middle

# Get Started

What is the difference between parametric or non-parametric?

---
# Bootstrap

- **Definition**: It usually refers to a self-starting process that is to proceed withot external input.
- Applied to statistics: We sample with replace from the sample
---
# Bootstrap
Bootstrap is a desirable approach when:
- **the distribution of a statistic is unknown or complicated**.
- **Reason**: bootstrap is a non-parametric and does not ask foor specific distributions.
- **the sample size is too small to draw a valid inference.**
- **Reason**: it is a resampling method with replacement and recreates any number of resamples.
---
# Let's breal down "bootstrap"

Bootstrap breaks down into the following steps:
- decide how many bootstrap samples to perform.
- what is the sample size?
- for each bootstrap sample:
- calculate the mean of the calculated sample statistics.

---
# Bootstrapping Illustration in R

```{r eval=FALSE, tidy=FALSE}
n <- 20
s <- rnorm(n = n,
           mean = 5.5,
           sd = 1.4)

library(bootstrap)
library(ggplot2)
bp_s <- bootstrap::bootstrap(s, 1000, var)
str(bp_s)
```

---

class: center, middle 

Other Sampling Approaches 
---
# Jackknife
**It is a leave-one-out procedure.**

---
# K-cross Validation

---
# Permutation 