---
title: "Nominal Dependent Variables"
subtitle: "Week 19"
author: "Winnie Xia"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
style_solarized_light()
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE, fig.align = 'center', out.width = "50%")
```

# Nominal Data in Political Science

- If outcome or dependent variable is binary, then use **logit or probit models**.

- If outcome or dependent variable is categorical but ordered (i.e low to high), then use **ordered logit or ordered probit models**
 - Agree; Neutral; Disagree
 - Low; Middle; High
 
- If outcome or dependent variable is categorical without any particular order, then use **multinomial logit/probit models**.

---
# Models for Nominal DVs
- Multinominal logit model: Series of binary logit models; similar to ordered logit. **Easiest to interpret**.
- Multinominal probit model: Series of binary probit models; similar to ordered probit. **Drops the IIA assumption**.
- Conditional logit model
- Joint multinominal and conditional logit model
- Mixed multinomial logit model
- Nested multinomial logit model
---
# R packages for Nominal families 
|  | nnet | mlogit | mnlogit | MNP |mclogit |
|--------| --- | --- | --- | --- | --- |
| Multinominal logit model | yes | yes | yes | no | yes |
| Multinominal probit model | no | yes | no | yes | no |
| Conditional logit model | no | yes| yes | no | yes |
| Joint multinominal and conditional logit model | no | yes | no | no | yes |
| Mixed multinomial logit model | no | yes | yes | no | no |
| Mixed multinomial logit model | no | no | yes | no | no |

---
# Multinomial Logit Model in R

```{r}
library(foreign)
library(texreg)
dat <- read.dta("https://stats.idre.ucla.edu/stat/data/hsbdemo.dta")
head(dat)
```

---
# Multinomial Logit Model in R

```{r}
library(nnet)
dat$prog2 <- relevel(dat$prog, ref = "academic")
m1 <- multinom(prog2 ~ ses + write, data = dat)
# prog, program type: general academic, and vocation
# ses, social economic status, categorical
# write, continuous, 
```
---
# Multinomial Logit Model in R

**Coefficients are log odds-ratios of the respective category over the reference category.**

```{r}
screenreg(m1, single.row = TRUE, beside = TRUE)
```

---
# Multinomial Logit Interpretation
```{r}
z <- summary(m1)$coefficients/summary(m1)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1))*2
data.frame(p)
```
---
# Interpretation
```{r}
data.frame(exp(coef(m1)))
```
- All else being equal, a one unit increase in **writing** decreases the relative odds of being in **general program** over **academic program** by approximately 0.94 %, and decreases the relative odds of **vocational program** over **academic program** by 0.89 %.

---
# Interpretation
- Again, we recommend, to use **predicted probabilities**.
```{r}
head(fitted(m1))
```
---
# Interpretation
- Then, we create a scenario to plot the predicted probabilities.
```{r}
library(reshape2)
newdat <- data.frame(ses = rep(c("low", "middle", "high"), 
                               each = 41),
                     write = rep(c(30:70), 3))

## store the predicted probabilities for each value of ses and write
pre <- cbind(newdat, predict(m1, newdata = newdat, type = "probs", se = TRUE))
## melt data set to long for ggplot2
lpp <- melt(pre, id.vars = c("ses", "write"), value.name = "probability")
```
- Here, we are looking at the averaged predicted probabilities is to look at the averaged predicted probabilities for different values of the continuous predictor variable **write** within each level of **ses**.
---
# Plot 
```{r }
library(ggplot2)
ggplot(lpp, aes(x = write, y = probability, colour = ses)) + 
  geom_line() + facet_grid(variable ~ ., scales = "free") +
  theme_minimal()
```

---
# Independence of Irrelevant Alternatives (IIA)
- IIA assumption states that the addition of a new category does not change the odds-ratios of the existing alternatives.
- In other words, the alternatives need to be **"distinct and weighed independently in the eyes of each decision maker"**.

---
# Test IIA assumption
- The Hausman-McFadden Test
```{r}
library(mlogit)
#mlogit::hmftest(prog2 ~ 0 | ses + write, data = dat,
                #alt.subset = c("general",
                               #"vocation"))
```
