---
title: "Nominal Dependent Variables"
subtitle: "Week 19"
author: "Winnie Xia"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
style_solarized_light()
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)
```

# Nominal Data in Political Science

- If outcome or dependent variable is binary, then use **logit or probit models**.

- If outcome or dependent variable is categorical but ordered (i.e low to high), then use **ordered logit or ordered probit models**
 - Agree; Neutral; Disagree
 - Low; Middle; High
 
- If outcome or dependent variable is categorical without any particular order, then use **multinomial logit/probit models**.

---
# Models for Nominal DVs
- Multinominal logit model: Series of binary logit models; similar to ordered logit. **Easiest to interpret**.
- Multinominal probit model: Series of binary probit models; similar to ordered probit. **Drops the IIA assumption**.
- Conditional logit model
- Joint multinominal and conditional logit model
- Mixed multinomial logit model
- Nested multinomial logit model
---
# R packages for Nominal families 
|  | nnet | mlogit | mnlogit | MNP |mclogit |
|--------| --- | --- | --- | --- | --- |
| Multinominal logit model | yes | yes | yes | no | yes |
| Multinominal probit model | no | yes | no | yes | no |
| Conditional logit model | no | yes| yes | no | yes |
| Joint multinominal and conditional logit model | no | yes | no | no | yes |
| Mixed multinomial logit model | no | yes | yes | no | no |
| Mixed multinomial logit model | no | no | yes | no | no |

---
# Multinomial Logit Model in R

```{r}
library(foreign)
library(texreg)
library(carData)
data(WVS)
WVS$poverty <- as.factor(as.character(WVS$poverty))
```

---
# Multinomial Logit Model in R

```{r}
library(nnet)
m1 <- multinom(poverty ~ religion + degree + country + 
                 age + gender, data = WVS)
```
---
# Multinomial Logit Model in R

**Coefficients are log odds-ratios of the respective category over the reference category.**

```{r}
screenreg(m1, single.row = TRUE, beside = TRUE)
```

---
# Multinomial Logit Interpretation
```{r}
z <- summary(m1)$coefficients/summary(m1)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1))*2
data.frame(t(p))
```
---
# Interpretation
```{r}
data.frame(t(exp(coef(m1))))
```
- All else being equal, every additional year of age decreases the relative odds of selecting **Too little** over **About Right** by approximately 1%, and increase the relative odds of **Too Much** over **About Right** by 0.7%.

---
# Interpretation
- Again, we recommend, to use **predicted probabilities**.
- Create a typical scenario and then plot them.
---
# Independence of Irrelevant Alternatives (IIA)
- IIA assumption states that the addition of a new category does not change the odds-ratios of the existing alternatives.
- In other words, the alternatives need to be **"distinct and weighed independently in the eyes of each decision maker"**.

---
# Test IIA assumption
- The Hausman-McFadden Test
```{r}

```


---
# The Multinomial Probit Model