---
title: "Limited Dependent Variables"
subtitle: "Week 20"
author: "Winnie Xia"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
# renderthis::to_pdf("slides.Rmd")

```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE, fig.align = 'center', out.width = "50%")
library(foreign)
library(tidyverse)
library(texreg)
```
# Limted Dependent Variable
- What is a limited dependent variable?
**Limited dependent variables** are dependent variables that have limited ranges, usually either **discontinuous** or **range bounded**.

---
# Censoring and Truncation
- Censoring is when the limit observations are in the sample.
- Truncation is when the observations are not in the sample.

---
# Truncated Regression
- Truncated regression addresses the bias introduced when using OLS regression with truncated data.
---
# Truncated Regression in R
```{r}
library(truncreg)
dat <- read.dta("https://stats.idre.ucla.edu/stat/data/truncreg.dta")
ggplot(dat, aes(achiv, fill = prog)) +
  geom_histogram(binwidth=3) +
  theme_minimal()
```
---
# Truncated Regression in R
```{r}
m <- truncreg(achiv ~ langscore + prog, data = dat, 
              point = 40, direction = "left")
screenreg(m)
```
---
# Tobit Model
The tobit model, also called a censored regression model, is designed to estimate linear relationships between variables when there is either left- or right-censoring in the dependent variable (also known as censoring from below and above, respectively).
---
# Tobit Model in R
```{r}
library(VGAM)
acad_apt = read_csv("https://stats.idre.ucla.edu/stat/data/tobit.csv") %>%
  mutate(prog = factor(prog, labels = c('acad', 'general', 'vocational')))
```

---
# Tobit Model in R
```{r}
m2 <- vglm(apt ~ read + math + prog, tobit(Upper = 800), data = acad_apt)
screenreg(m2)
```
---
# Heckit Model
- Heckit model can be used when there is incidental truncation, i.e., there is variable that explains whether we have observations for an outcome variable.
- **Example**:
  - wage for those who have jobs, another variable that indicates whether someone has a job;
  - grades for those who go to universities, another variable that indicates whether someone goes to university or not.
- Two-Stage least squares, **a statistical technique that is used in the analysis of structural equations.**
---
# Heckit Model in R
