---
title: "Limited Dependent Variables"
subtitle: "Week 20"
author: "Winnie Xia"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)

style_mono_dark(base_color = "#cbf7ed")
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE, fig.align = 'center', out.width = "50%")
```
# Limted Dependent Variable
- What is a limited dependent variable?
**Limited dependent variables** are dependent variables that have limited ranges, usually either **discontinuous** or **range bounded**.

---
# Censoring and Truncation
- Censoring is when the limit observations are in the sample.
- Truncation is when the observations are not in the sample.

---
# Truncated Regression

---
# Truncated Regression in R
```{r}
library(truncreg)
```


---
# Tobit Model

---
# Tobit Model in R
```{r}
library(tidyverse)


acad_apt = read_csv("https://stats.idre.ucla.edu/stat/data/tobit.csv") %>%
  mutate(prog = factor(prog, labels = c('acad', 'general', 'vocational')))
```


---
# Heckit Model

---
# Heckit Model in R



