---
title: "Counts and Proportions"
subtitle: "Week 21"
author: "Winnie Xia"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#23395b",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
# renderthis::to_pdf("slides.Rmd")

```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE, fig.align = 'center', out.width = "50%")
library(foreign)
library(tidyverse)
library(texreg)
```

# Negative Binomial Model
* The negative binomial model introduces a heterogeneity parameter to rectify the problem of over-dispersion for count data. 

---
# Negative Binomial in R
We use the `glm.nb` function from the **MASS** package to estimate a negative binomial regression.
---
# Hurdle Model
* Hurdle models are applied to situations in which target data has relatively many of one value, usually zero, to go along with the other observed values. They are two-part models, a logistic model for whether an observation is zero or not, and a count model for the other part. 
* The key distinction from the usual ‘zero-inflated’ count models, is that the count distribution does not contribute to the excess zeros. 

---
# Hurdel Model in R

---
# Zero-inflated Poisson Models
* It is used for modeling count variables with excessive zeros and it is usually for over-dispersed count outcome variables.


---
# Zero-truncated Negative Binomial Model 
* It is used to model count date for which **the value zero cannot occur** and **when there is evidence of over dispersion**

* Examples:
 + A study of the number of journal articles published by tenured faculty as a function of discipline (fine arts, science, social science, humanities, medical, etc). To get tenure faculty must publish, i.e., there are no tenured faculty with zero publications.
 + A study of length of hospital stay, in days, as a function of age, kind of health insurance and whether or not the patient died while in the hospital. Length of hospital stay is recorded as a minimum of at least one day.
---
# Zero-truncated Negative Binomial Model in R
