---
title: "Ordinal Dependent Variables"
subtitle: "Week 18"
author: "Winnie Xia"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
style_duo(primary_color = "#1F4257", secondary_color = "#F97B64")
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)
```

# Ordered logit and probit regression

- If outcome or dependent variable is binary, then use **logit or probit models**.

- If outcome or dependent variable is categorical but ordered (i.e low to high), then use **ordered logit or ordered probit models**
 - Agree; Neutral; Disagree
 - Low; Middle; High
 
- If outcome or dependent variable is categorical without any particular order, then use **multinomial logit models**.

---
# Ordered logit and probit in R
```{r}
library(foreign)
dat <- read.dta("https://stats.idre.ucla.edu/stat/data/ologit.dta")
head(dat)
```
---
# Ordered logit and probit in R
Below we use the `polr` function from the MASS package to estimate an ordered logistic regression model. 
```{r, message=FALSE}
library(MASS) 
library(texreg)
m1 <- polr(apply ~ pared + public + gpa, data = dat, 
           method = "logistic", Hess=TRUE)

m2 <- polr(apply ~ pared + public + gpa, data = dat, 
           method = "probit", Hess=TRUE)

m3 <- glm(as.numeric(apply) ~ pared + public + gpa,
          data = dat)

```
---
# Ordered logit and probit in R
```{r , echo=FALSE}
screenreg(list(m1, m2, m3))
```
---
# Interpretation
```{r }
summary(m1)
```
---
# Interpretation
We often interpret models by calculating marginal effects for an ordered choice model and their standard errors.
```{r message=FALSE}
library(erer)
marg <- ocME(m1)
marg
```
---
# Interpretation
```{r message=FALSE}
marg$out # get t and p-values

```
---
# The Parallel Regression Assumption
- An alternative way of conceptualizing the ordered logit/probit model is as the constrained estimation of a system of logit/probit models.

- Parallel regressions describe how the impact of covariates may shift the predicted probability curves to the right ot left, but **does not change the basic slope of these curves across any two categories.**
---
# Assessing the assumption

- Fit separate binary logit or probit models for the different cut-off levels and see if the slopes are similar.

- The Brant test: A p-value of less than 0.05 on this test —particularly on the Omnibus plus at least one of the variables— should be interpreted as a failure of the proportional odds assumption.

```{r }
library(brant)
brant(m1)
```

---
# Cumulative Link Model
```{r}
library(ordinal)
m4 <- clm(apply ~ pared + public + gpa, data = dat, 
           link = "logit")

```

---
# Cumulative Link Model
```{r}
summary(m4)
```

